% Features to discuss in-depth:

SymPy has an extensive feature set that encompasses too much to cover in-depth
here. Bedrock areas, such as calculus, receive their own subsections below.
Table~\ref{features-table} gives a compact listing of all major capabilities
present in the SymPy codebase. This gives a sampling from the breadth of
topics and application domains that SymPy services. Note that unless stated
otherwise, all features noted in Table~\ref{features-table} are symbolic in
nature. Numeric features are discussed in Section~\ref{sec:numerics}.


\begin{longtable}[htbc]{p{0.25\linewidth}p{0.68\linewidth}}
\caption{SymPy Features and Descriptions\label{features-table}}\\
\toprule
\textbf{Feature} & \textbf{Description} \\
\midrule
Calculus & Algorithms for computing derivatives, integrals, and limits.\\

Category Theory & Representation of objects, morphisms, and diagrams. Tools
for drawing diagrams with Xy-pic.\\

Code Generation & Enables generation of compilable and executable code in a
variety of different programming languages directly from expressions. Target
languages include C, Fortran, Julia, JavaScript,
Mathematica, Matlab and Octave, Python, and Theano.\\

Combinatorics \& Group Theory & Implements permutations, combinations,
partitions, subsets, various permutation groups (such as polyhedral, Rubik,
symmetric, and others), Gray codes~\cite{Nijenhuis1978combinatorial},
and Prufer sequences~\cite{biggs1976graph}.\\

Concrete Math & Summation, products, tools for determining whether summation
and product expressions are convergent, absolutely convergent, hypergeometric,
and
other properties. May also compute Gosper's normal form~\cite{petkovvsek1996bak} for two univariate polynomials.\\

Cryptography & Represents block and stream ciphers, including shift, Affine,
substitution, Vigenere's, Hill's, bifid, RSA, Kid RSA,
linear-feedback shift registers, and Elgamal encryption\\

Differential Geometry & Classes to represent manifolds, metrics, tensor
products, and coordinate systems.\\
% TODO: Someone verify that this is a good summary of the diffgeom module

Geometry & Allows the creation of 2D geometrical entities, such as lines and
circles. Enables queries on these entities, such as asking the area of an
ellipse, checking for collinearity of a set of
points, or finding the intersection between two lines.\\

Lie Algebras & Represents Lie algebras and root systems.\\

Logic & boolean expression, equivalence testing, satisfiability, normal
forms.\\

Matrices & Tools for creating matrices of symbols and expressions. This is
capable of both sparse and dense representations and performing
symbolic linear algebraic operations (e.g., inversion and factorization).\\

Matrix Expressions & Matrices with symbolic dimensions (unspecified entries).
Block matrices.\\

Number Theory & prime number generation, primality testing, integer
factorization, continued fractions, Egyptian fractions, modular arithmetic,
quadratic residues, partitions, binomial and multinomial coefficients,
prime number tools, integer factorization. \\

Plotting & Hooks for visualizing expressions via matplotlib~\cite{Hunter:2007}
or as text drawings when lacking a graphical back-end. 2D function plotting,
3D function
plotting, and 2D implicit function plotting are supported.\\

Polynomials & Computes polynomial algebras over various coefficient domains.
Functionality ranges from the simple (e.g., polynomial division) to the
advanced (e.g., Gr\"obner bases~\cite{adams1994introduction} and multivariate
factorization over algebraic number domains).\\

Printing & Functions for printing SymPy expressions in the terminal with ASCII
or Unicode characters, and converting SymPy expressions to \LaTeX{} and
MathML.\\

Series & Implements series expansion, sequences, and limit of sequences.
This includes special series, such as Fourier and formal power series.\\

Sets & Representations of empty, finite, and infinite sets. This includes
special sets such as for all natural, integer, and complex numbers. Operations
on sets such as union, intersection, Cartesian product, and building sets from
other sets.\\

Simplification & Functions for manipulating and simplifying expressions.
Includes algorithms for simplifying hypergeometric functions, trigonometric
expressions, rational functions, combinatorial functions, square root
denesting, and common subexpression elimination.\\

Solvers & Functions for symbolically solving equations algebraically, systems
of equations, both linear and non-linear, inequalities, ordinary differential
equations, partial differential equations, Diophantine equations, and
recurrence relations.\\

Special Functions & Implements a number of well known special functions,
including Dirac delta, Gamma, Beta, Gauss error functions, Fresnel integrals,
Exponential integrals, Logarithmic integrals, Trigonometric integrals, Bessel,
Hankel, Airy, B-spline, Riemann Zeta, Dirichlet eta, polylogarithm, Lerch
transcendent, hypergeometric, elliptic integrals, Mathieu, Jacobi polynomials,
Gegenbauer polynomial, Chebyshev polynomial, Legendre polynomial, Hermite
polynomial, Laguerre polynomial, and
spherical harmonic functions.\\

Statistics & Support for a random variable type as well as the ability to
declare this variable from prebuilt distribution functions such as
Normal, Exponential, Coin, Die, and other custom distributions.\\

Tensors & Symbolic manipulation of indexed objects.\\

Vectors & Provides basic vector math and differential calculus with respect
to 3D Cartesian coordinate systems.\\
\bottomrule

\end{longtable}

\subsection{Simplification}

\input{simplification}

\subsection{Calculus}
\label{sec:calculus}
\input{calculus}

\subsection{Printers}

\input{printers.tex}

% Solvers (regular equations, maybe also mention other types of solvers like ODEs/recurrence/Diophantine)
\subsection{Solvers}
\input{solvers.tex}

% Matrices (worth including to stress that they are symbolic)
\subsection{Matrices}

\input{matrices}
